\textcolor{red}{Simple} connectome inference using partial correlation from imaging calcium signal

### 2014-06-16 article.tex
fix ambiguity with separated graph
re-wrote directivity section

Since all neurons are assumed to be connected to at least one neuron, we thus
have by definition $X_i \not\perp X_j$ \textcolor{red}{AS: Je ne suis pas sûr
de comprendre.}. However,  two random variables $A$ and $B$ might be
dependent taken alone ($A \not\perp B$), but not once we observe the set  $C$
of all other variables ($A \perp B | C$). The conditioning over all other
random variables $X^{-i,j}$ is thus necessary to avoid predicting an edge when
two neurons are not directly connected.

@@

\subsection{Causal discovery methodology}
% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.}.

By ignoring the directivity, we systematically make a true positive and a
false positive for direct link.  Hoping to orientate
correctly the most obvious directed links, we attempt to retrieve some causal
information, i.e. the directivity of the links, by computing and stacking with
our symmetric adjacency matrix a matrix of pairwise activation heuristics.
This method computes an anti-symmetric activation score $Z$ based the
variation of fluorescence signal of a neuron $j$ due to a neuron $i$ over all
time steps:
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_{t+1,j} - x_{t, i} \in \left[f1, f2 \right]) -
    \mathbb{1}(x_{t+1,i} - x_{t, j} \in \left[f1, f2 \right])
\]
where $f_1$ and $f_2$ are parameters of the method.

>>>>

Since all neurons are assumed to form a connected graph, we
have by construction $X_i \not\perp X_j$. However, two random variables $A$ and $B$ might be
dependent taken alone ($A \not\perp B$), but not once we observe the set  $C$
of all other variables ($A \perp B | C$). The conditioning over all other
random variables $X^{-i,j}$ is thus necessary to avoid predicting an edge when
two neurons are not directly connected.

@@

% Introduce a custom solution to make the Y matrix asymmetric
% Introduce directivity
Since partial correlation is a symmetric measure, the causal mechanism behind the
interaction ($A$ directly causes $B$, $B$ directly causes $A$ or both) can not
be directly inferred\footnote{Also note that two other causal mechanisms might be
implied while having a non-zero partial coefficient: (i) a pair of variables
is induced by a common (hidden) variable; (ii) $A$ (respectively $B$) is
conditionally correlated to a hidden variable affecting $B$ (resp. $A$)
\cite{de2004discovery}.}.
In order to retrieve causal information, we developed an
heuristics based on the following observation. The rise of fluorescence
indicates the activation of a neuron $i$. If a neuron $j$ have also
an increased of fluorescence after a slight delay, this could have been
caused by the neuron $i$. Thus, we count whenever
such events occurred between consecutive time steps
\[
\hat{Z}: z_{ij} = \sum_{t=1}^{T - 1}
    \mathbb{1}(x_j^{t+1} - x_i^t \in \left[f_1, f_2 \right]) -
    \mathbb{1}(x_i^{t+1} - x_j^t \in \left[f_1, f_2 \right]), \forall i, j \in V
\]
where $f_1$ and $f_2$ are parameters of the method.
Finally, the weighted average of correlation matrices is combined with $Z$ such
that $Z$ has a contribution of $0.3\%$ to the connectivity score matrix.
